---
title: "Causal_inference_NatFood"
output: html_document
date: "2023-06-24"
author: Liya Weldegebriel
---
Description: 
- imports daily rainfall, cultivation status and number of conflict incidences data using 10 by 10 km grid data
- calculates two-way fixed effect estimates of the impact of the number of conflict incidence on area of well-cultivated land with rainfall characterstic covariates


```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
library(reshape2)
library(fixest)

```

#Import data

```{r}
files.location <- paste0('UPDATE WITH YOUR FILE LOCATION') #Update with local file location for all data

```

##Importing daily rainfall 2019 to 2021 

```{r}
setwd(files.location) 
read.csv("Precip_daily_Tigray_19to21_529grids.csv")-> daily.precip
```



##Importing Cultivatetion status datafrom pre-war (2019/2020) and post-war (2021) periods

```{r}
setwd(files.location) 

read.csv("CultArea_cropgrass_1200_10k_2021.csv")%>%
  mutate(., cult_area = X1, year = 2021) %>%
  select(., c(id,cult_area,year)) ->stat_21

read.csv("CultArea_cropgrass_1200_10k_2020.csv")%>%
  mutate(., cult_area = X1, year = 2020)%>%
  select(., c(id,cult_area, year))->stat_20

rbind(stat_21, stat_20)%>%
  mutate(., cult_area = cult_area/1E6)-> cult_stat

head(cult_stat)
```

##Importing conflict incidence data

```{r}
setwd(files.location) 

read.csv("CI_count_10k.csv")%>%
  mutate(., grid_size = "ten_k")%>%
  group_by(id, grid_size)%>%
  summarise(count = sum(n_CI))-> grid_CI_count

grid_CI_count%>%
  mutate(., CI = count)%>%
  select(., c("id","CI"))-> CI
head(CI) 
```


```{r}
setwd(files.location) 

read.csv("CI_count_10k.csv")-> CI_raw


with(CI_raw, 
     ifelse(month == "2020" ,'early',
     ifelse(month == "jan" ,'early',
     ifelse(month == "feb" ,'prep', 
     ifelse(month == "march" ,'prep',
     ifelse(month == "april" ,'prep',
     ifelse(month == "may" ,'prep',
     ifelse(month == "june" ,'growing',
     ifelse(month == "july" ,'growing',
     ifelse(month == "aug" ,'growing',
     ifelse(month == "sept" ,'growing',
     ifelse(month == "oct" ,'growing', "NA"))))))))))))-> CI_raw$time.group

CI_raw%>%
  filter(., time.group == "prep")%>%
  mutate(., CI = n_CI)%>%
  select(., c(id,CI))-> CI_prep

CI_raw%>%
  filter(., time.group == "early")%>%
  mutate(., CI = n_CI)%>%
  select(., c(id,CI))-> CI_early

CI_raw%>%
  filter(., time.group == "prep" | time.group == "growing" )%>%
  mutate(., CI = n_CI)%>%
  select(., c(id,CI))-> CI_prep_growing

CI_raw%>%
  filter(., time.group == "growing" )%>%
  mutate(., CI = n_CI)%>%
  select(., c(id,CI))-> CI_growing

CI_prep
CI_early
CI_prep_growing
CI_growing
```



#Rainfall characterstics

```{r}
require(lubridate)

daily.precip%>%
  mutate( day = date, Rain = precipitation )%>%
  mutate('date' = make_date(year = year, month = month, day = date))%>%
  select(., c("date","day","month","year", "id","Rain")) -> rf
head(rf)

```



```{r}
rf%>%
  filter(month %in% unique(.$month)[c(6,7,8,9)]) -> Kiremt_rf #June to Sept

head(Kiremt_rf)
```


##Rainfall Frequency


Initializing a data frame
```{r}
frequency_initial <- data.frame(Rain_day=numeric(),
                           No_Rain_day=numeric(),
                           NA_s=numeric(),
                           frequency_RF=numeric(),
                           year = integer(),
                           id = integer(),
                           stringsAsFactors=FALSE)

temp_Rain <- data.frame(Rain_day=numeric(),
                           No_Rain_day=numeric(),
                           NA_s=numeric(),
                           frequency_RF=numeric(),
                           year = integer(),
                           id = integer(),
                        
                           stringsAsFactors=FALSE)

```

Rainfall frequency function 
```{r}
freq <- function(Stations_Data) {
  
  yr_id <- c(2019,2020,2021)
  id <- c(0:528)
  
  for(i in yr_id){  
    for(j in id){
      
      Stations_Data%>%
        filter(id == j)%>%
        filter(year == i) ->V
      
        length(which(V$Rain!=0))-> frequency_initial[1,"Rain_day"]
        length(which(V$Rain==0)) -> frequency_initial[1,"No_Rain_day"]
        length(which(is.na(V$Rain))) -> frequency_initial[1,"NA_s"]
      
      frequency_initial$year <- i
      frequency_initial$id <- j
      
   
      temp_Rain <- rbind(temp_Rain,frequency_initial)
    }
  }
  return(temp_Rain)
}
```


```{r}
freq(Stations_Data = Kiremt_rf)-> Rainfall_Frequency_Kiremt

Rainfall_Frequency_Kiremt$Rain_day/Rainfall_Frequency_Kiremt$No_Rain_day-> Rainfall_Frequency_Kiremt$frequency_RF 
Rainfall_Frequency_Kiremt

```
```{r}
Rainfall_Frequency_Kiremt%>%
  group_by(year)%>%
  summarise_at(c("frequency_RF"), mean, na.rm=TRUE)-> meanRFfreq

meanRFfreq$year <- as.factor(meanRFfreq$year)

```



##Mean rainfall

```{r}
mean_initials <- data.frame(Mean_Kiremt=numeric(),
                           year = integer(),
                           id = integer(),
                           stringsAsFactors=FALSE )

temp_MeanRains <- data.frame(Mean_Kiremt=numeric(),
                           year = integer(),
                           id = integer(),
                           stringsAsFactors=FALSE )

Mean_rain <- function(Stations_Data) {
  
  yr_id <- c(2019,2020,2021)
  id <- c(0:529)
  
  for(i in yr_id){  
    for(j in id){
      
      Stations_Data%>%
        filter(id == j)%>%
        filter(year == i)->V_kiremt
      
    V_kiremt%>%
      filter(., Rain!=0) -> V_kiremt
          
      mean(V_kiremt$Rain, na.rm= TRUE) ->mean_initials[1,"Mean_Kiremt"]
          
      
        
      mean_initials$year <- i
      mean_initials$id <- j
      
   
      temp_MeanRains <- rbind(temp_MeanRains,mean_initials)
    }
  }
  return(temp_MeanRains)
}

```

Finding Means
```{r}
Mean_rain(Stations_Data =Kiremt_rf )-> Rainfall_Mean
Rainfall_Mean
```


##Cumulative rainfall

```{r}
Kiremt_rf%>%
  dplyr::select(Rain, id, year)%>%
  group_by(id, year)%>%
  summarize(Rain_Cum=sum(Rain,na.rm=TRUE))-> Rainfall_cum_kiremt
Rainfall_cum_kiremt

rf%>%
  dplyr::select(Rain, id, year)%>%
  group_by(id, year)%>%
  summarize(Rain_Cum=sum(Rain,na.rm=TRUE))-> Rainfall_cum_annual
Rainfall_cum_annual


Rainfall_cum_annual%>%
  group_by(id) %>%
  summarise_at(vars(-year), funs(mean(., na.rm=TRUE))) ->Mean_Annual_rain

head(Mean_Annual_rain)
```

##Summary of cumulative, mean and frequency of kiremt rainfall

```{r}
head(Rainfall_cum_kiremt)
head(Rainfall_Mean)
head(Rainfall_Frequency_Kiremt)

```

Combining RF characteristics
```{r}
Rainfall_Frequency_Kiremt %>%
  select(., c(frequency_RF, year, id))%>%
  merge(., Rainfall_Mean, by = c("year","id"))%>%
  merge(.,Rainfall_cum_kiremt, by = c("year","id") ) -> rf_summary

head(rf_summary)
```
##RF summary

```{r}
rf_summary$year = ifelse(rf_summary$year <= 2020, 2020, 2021)

rf_summary%>%
  group_by(id,year)%>%
    summarise_at(c("frequency_RF", "Mean_Kiremt", "Rain_Cum"), mean, na.rm=TRUE)-> rf.data

head(rf.data)

```


#Causal inference analysis via TWFE

```{r}
twfe_function <- function(rf.data, cult.data, CI.data){
  
  
rf.data%>%
   merge(., cult.data, by = c("id", "year"))%>%
   merge(., CI.data, by = "id") -> rf_twfe

rf_twfe$CI = ifelse(rf_twfe$year == 2020, 0, rf_twfe$CI) #replace all year 2020 CI with 0


rf_twfe$id <- as.factor(rf_twfe$id)
rf_twfe$year <- as.factor(rf_twfe$year)



return(summary(feols(cult_area ~ CI + Rain_Cum + frequency_RF + Mean_Kiremt | id + year, data = rf_twfe)))
  
  
}
```



##Using All CIs
```{r}
twfe_function(rf.data, cult_stat, CI)
```
##Using Growing CIs
```{r}
twfe_function(rf.data, cult_stat, CI_growing)
```

##Using Prep CIs
```{r}
twfe_function(rf.data, cult_stat, CI_prep)
```

##Using Early CIs
```{r}
twfe_function(rf.data, cult_stat, CI_early)
```

